Stock enhancement undermines community stability

### 2022-02-01 rmd/manuscript.Rmd
Update manuscript.Rmd

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Captive breeding is common to enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. Although there is growing recognition that releases of captive-bred organisms (hereafter, "stock enhancement") entail ecological risks (e.g., the accumulation of deleterious alleles) [@arakiGeneticEffectsCaptive2007; @krkosekDecliningWildSalmon2007], this method is still pervasive in conservation [@snyderLimitationsCaptiveBreeding1996] and natural resource management [@laikreCompromisingGeneticDiversity2010]. In fisheries, for example, billions of hatchery individuals are released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks.

@@

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Therefore, the model accounts for the fitness difference of captive-bred individuals due to genetic effects and/or plasticity [@weberInteractionsHatcheryWild2003; @snyderLimitationsCaptiveBreeding1996; @arakiHatcheryStockingHelp2010] when considering the reproductive contribution of released individuals to the next generation. Without loss of generality, we assumed constant $K_i$ ($K_i = K$) and $R_t$ ($R_t = R$) across species and time, respectively. We ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates) under each of `r nrow(param_set)` simulation scenarios that cover a range of ecological contexts (see **Materials and Methods**). The last `r unique(sim_result$n_timestep)` time steps were used to obtain the following summary statistics of the total community density $\sum_i^SN_i$: the coefficient of variation (CV), the number of species persist (defined as $N_i > 0.01$ at $t=1600$), the temporal mean ($\mu$), and the temporal SD ($\sigma$). We also calculated the temporal mean and SD for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer mechanisms that underlie the community-level response.

When carrying capacity is small ($K=100$), our model predicted a destabilizing effect of stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (**Figure 1A**). This pattern stemmed mainly from the reduced mean of total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). For enhanced species, released individuals induced overcompensation, in which the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals. Therefore, the enhanced species reached the lower long-term average of population density with stocking, confirming previous theoretical predictions [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species with fewer persisting species at high levels of stock enhancement (**Figure 1A**). Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density resulted in the increased CV. These patterns were qualitatively similar across most ecological scenarios (**Figure S1-6**).

@@

Stock enhancement, however, had little influence on community dynamics when carrying capacity was sufficiently large ($K=500$; **Figures S7-S12**). In particular, stock enhancement increased the enhanced species with a low population growth rate ($r_1=0.5$; **Figures S7 and S8**). This result may explain why some of the best evidence for successful stock enhancement comes from long-lived endangered species [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. More importantly, the contrasting community response at different carrying capacities provides deeper insights. One of the core motivations for stock enhancement is to mitigate declining trends of natural resources due to human impacts, such as habitat loss and fragmentation [@chaseEcosystemDecayExacerbates2020]. Ironically, our results suggest that stock enhancement will not bring desired outcomes unless we resolve root causes that compromise environmental capacity first.

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon *Oncorhynchus masou masou* (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. Masu salmon is one of Japan's most valuable freshwater resources, and over 10 million hatchery fish are released annually across the nation [@kitadaLessonsJapanMarine2020]. In the protected watersheds (**Figure 1C and 1D**), the Hokkaido government led the standardized monitoring of fish communities over the past six decades, along with the official stocking records. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. In this study, we used the long-term community data (1999-2019) at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` protected watersheds (see Methods for selection criteria). Using hierarchical Bayesian models, we quantified the effect of stock enhancement on community dynamics while accounting for potential confounding effects of climates and local abiotic factors.

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with greater numbers of fish released (**Figure 1B**). Our analysis strongly supports the positive relationship between the CV and the number of releases (**Figure 2**), in which the probability of the regression coefficient being positive was `r op(df_beta$all$prob_positive[df_beta$all$response == "cv"])` (**Table S8**). This pattern was associated with the reduced long-term average of total community density and fewer species richness (**Figure 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figure 1B and 2;** see **Tables S8-S10** for full statistics). In the meantime, the SDs had vague relationships with stock enhancement (**Figure 1B and 2**).

Competition with wild masu salmon and other fish species is a likely mechanism, as our theory assumes. Field experiments confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage, typically one year [@hasegawaEffectsHatcheryChum2014; @hasegawaWildMasuSalmon2018]. Hatchery salmons are larger and more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @weberInteractionsHatcheryWild2003]. Importantly, socioeconomic factors control the number of releases (e.g., budget allocation). As such, released fish are "excessive" given the density-dependent regulation of recipient communities and may cause resource competition that would otherwise not exist.

The reduced fitness of hatchery masu salmon may also play a role in the observed response. Like other salmonids, ocean-migrating adults of hatchery masu salmon show lower return rates to the spawning river [@sahashiWildGenesBoostinpress]. Hatchery salmons are also more susceptible to predation than wild ones [@weberInteractionsHatcheryWild2003]. The fitness disadvantage of captive-bred fish may add to negative density dependence to influence community stability.

>>>>

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Captive breeding is common to enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. Although there is growing recognition that releases of captive-bred organisms (hereafter, "stock enhancement") entail ecological risks (e.g., the accumulation of deleterious alleles) [@arakiGeneticEffectsCaptive2007; @krkosekDecliningWildSalmon2007], this method is still pervasive in conservation [@snyderLimitationsCaptiveBreeding1996] and natural resource management [@laikreCompromisingGeneticDiversity2010]. In fisheries, for example, billions of hatchery individuals are released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that demographic benefits of released individuals may exceed the associated risks.

@@

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Therefore, the model accounts for the fitness difference of captive-bred individuals due to genetic effects and/or plasticity [@weberInteractionsHatcheryWild2003; @snyderLimitationsCaptiveBreeding1996; @arakiHatcheryStockingHelp2010] when considering the reproductive contribution of released individuals to the next generation. Without loss of generality, we assumed constant $K_i$ ($K_i = K$) and $R_t$ ($R_t = R$) across species and time, respectively.

We ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates) under each of `r nrow(param_set)` simulation scenarios. These scenarios cover a range of ecological contexts, differing in the intrinsic growth of the enhanced species $r_1$, competition $\bar{\alpha}$, carrying capacity $K$, environmental stochasticity $\sigma_{\epsilon}$, and relative fitness $f_R$ (see **Materials and Methods**). The last `r unique(sim_result$n_timestep)` time steps were used to obtain the following summary statistics of the total community density $\sum_i^SN_i$: the coefficient of variation (CV), the number of species persist (defined as $N_i > 0.01$ at $t=1600$), the temporal mean ($\mu$), and the temporal SD ($\sigma$). We also calculated the temporal mean and SD for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer mechanisms that underlie the community-level response.

When carrying capacity is small ($K=100$), our model predicted a destabilizing effect of stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because stocking induced the negative competitive effect exceeding the demographic benefit of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of stock enhancement, resulting in fewer persisting species (**Figure 1A**). Combined, the long-term average of the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar in a range of ecological contexts (robust to the changes in $r_1$, $\bar{\alpha}$, $\sigma_{\epsilon}$, and $f_R$; **Figure S1-6**).

@@

Stock enhancement, however, had little influence on community dynamics when carrying capacity was sufficiently large ($K=500$; **Figures S7-S12**). In particular, stock enhancement increased the enhanced species with a low population growth rate ($r_1=0.5$; **Figures S7 and S8**). This result may explain why some of the best evidence for successful stock enhancement comes from long-lived endangered species [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. More importantly, the contrasting community response at different carrying capacities provides deeper insights. One of the core motivations for stock enhancement is to mitigate declining trends of natural resources due to human impacts, such as habitat loss [@chaseEcosystemDecayExacerbates2020]. Ironically, our results suggest that stock enhancement will not bring desired outcomes unless we resolve root causes that compromise environmental capacity first.

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of the stock enhancement of masu salmon *Oncorhynchus masou masou* (**Figure 1E**) on the long-term stability of stream fish communities in Hokkaido, Japan. Masu salmon is one of Japan's most valuable freshwater resources, and over 10 million hatchery fish are released annually [@kitadaLessonsJapanMarine2020]. In the protected watersheds (**Figure 1C and 1D**), a long-term program exists to monitor stream fish communities along with the official stocking records. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. We used the data from 1999 to 2019 at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` protected watersheds (see **Materials and Methods** for selection criteria). Using hierarchical Bayesian models, we quantified the effect of stock enhancement on community dynamics while accounting for potential confounding effects of climates and local abiotic factors.

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with intensive stocking (**Figure 1B**). Our analysis strongly supported the positive relationship between the CV and the number of releases (**Figure 2**), in which the probability of the regression coefficient being positive was `r op(df_beta$all$prob_positive[df_beta$all$response == "cv"])` (**Table S8**). This pattern was associated with the reduced long-term average of total community density and fewer species richness (**Figure 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figure 1B and 2;** see **Tables S8-S10** for full statistics). In the meantime, the SDs had vague relationships with stock enhancement (**Figure 1B and 2**).

Competition is a likely mechanism, as our theory assumes. Field experiments confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage, typically one year [@hasegawaEffectsHatcheryChum2014; @hasegawaWildMasuSalmon2018]. Hatchery salmons are larger and more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @weberInteractionsHatcheryWild2003]. Importantly, socioeconomic factors control the number of releases (e.g., budget allocation). As such, released fish are probably "excessive" given the density-dependent regulation of recipient communities and may cause resource competition that would otherwise not exist.

The reduced fitness of hatchery masu salmon may also play a role in the observed response. Like other salmonids, ocean-migrating adults of hatchery masu salmon show lower return rates to the spawning river [@sahashiWildGenesBoostinpress]. In addition, hatchery salmons are more susceptible to predation than wild ones [@weberInteractionsHatcheryWild2003]. The fitness disadvantage of hatchery fish may add to negative density dependence to influence community stability.
